Coherent optical association of a single molecule

### 2020-09-21 paper.tex
Update on Overleaf.

Such ultracold molecular association takes advantage of the much developed cooling and trapping techniques for atoms  as a starting point. To overcome the challenges of small wavefunction overlap and the large release of binding energy of converting atoms to deeply-bound molecules, a  two-step approach have been established to first associate atom pairs into weakly-bound molecules, and then transfer the molecules from one internal state  to a desired rovibrational and electronic state~\cite{Danzl2008, Ni2008,Lang2008, Takekoshi2014, Molony2014, Park2015, Guo2016, Kondov2019, Voges2020}.
So far, all of such association process  utilize a magnetic Feshbach scattering resonance and  apply to bialkali molecules. The only exception is Sr$_2$~\cite{Reinaudi2012,Stellmer2012} where narrow linewidth excited states are available and optical association can be driven. The requirement of a Feshbach resonance to enhance atom-to-molecule wavefunction overlap or the narrow excited state lines limits the generality of the association technique to more diverse molecular species.

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule using an optical Raman transfer via an electronic excited state, without the use of a Feshbach resonance nor a narrow excited state. Our scheme is based on a theoretically determined  vibrational state of the electronic excited state $c^3\Sigma^+$ that has the best transfer Rabi coupling to photon scattering ratio. Furthermore, to reduce technical requirements such as intensity stability, etc , we choose a initial and final state ... The choice of the states is general and can be applied to associate other molecular species or large molecules atom-by-atom.
%electronic the best initial hyperfine combination of the atoms to use and best single photon frequency that balances coupling/transfer efficiency and photon scattering. Using these parameters, we demonstrate coherent association of a electronic ground-state molecule from two atoms via an optical Raman transition in an optical tweezer.

@@

\begin{figure*}
  \includegraphics[height=4.5cm]{fig2.pdf}
  \caption{
    (A) Geometry and polarization of trap and Raman beam relative to the bias magnetic field.
    The tweezer and Raman beam is focused through an objective to define the location of the
    atoms and molecule.
    We use a bias B field of $B_0=8.8 G$ alone the tweezer polarization
    to define the quantization axis.
    As a result, the atoms experiences predominately $\pi$ polarization from the tweezer.
    (B) Molecule formation pulse sequence. The tweezer initially consists of only up leg power.
    When driving the Raman transition, the up leg power is smoothly ramped down and
    the down leg power ramped up over $10\mu s$ while maintaining the total power of the tweezer.
    This minimizes the heating on the atoms due to power fluctuation while maximizes the time
    with maximum Raman Rabi frequency when the up and down leg powers are equal.
    (C) Raman resonance from atomic state to molecular state, showing Fourier limited linewidth.
    \todo{states, time}
    (B) Rabi oscillation on resonance
    \label{f-raman}}
\end{figure*}

@@

While optical Raman transfer of atoms to molecules utilizing an electronic excited state has been shown previously\cite{Wynar2000,Rom2004}. Those demonstrations were  incoherent  where the signal shows resonance loss of population and the transferred molecules cannot be further utilized(need to improve). The challenge lies at  achieve high enough Raman Rabi frequency to drive this transition due to the wavefunction size mismatch with the excited molecular state.
 Previous experiments use weakly bound molecular excited states in the Raman transition to increase the Raman Rabi frequency. However, this choice of state also increases the scattering during the transfer process for two reasons.
Firstly, since for a typical molecular potential the spacing between states
decreases with increasing vibrational levels,
a weakly bound excited molecular state limits maximum usable detuning and therefore the effectiveness of scattering reduction by increasing the detuning for the Raman transition.
Secondly, due to size similarity, the weakly bound molecular ground state has a strong coupling with the atomic excited state which contributes significantly to the total scattering rate.
The rate of such scattering process is proportional to $1/\delta_{thresh}^2$,
where $\delta_{thresh}$ is the detuning from the dissociation threshold. (Should we also mention scattering of the atomic initial state too? This should affect the Raman process as well)
% This further increases the scattering rate for a weakly bond excited state.
Based on these considerations, we calculated both the Raman Rabi frequency and the scattering rate at different detunings (full result is shown in SM). As seen in figure (1....), despite a higher Raman Rabi frequency at the same detuning, using a weakly bound intermediate state results in a much higher scattering rate and ultimately a smaller Rabi frequency to scattering rate ratio as compared to deeply bound states. As a result, we use a deeply bound molecular excited state to drive the Raman transition. (mention the trend, also v=12 will work as similarly as v=0)

In addition to the intermediate state, the choice of the initial and final states are critical for both fundamental and technical reasons. A key difference in molecular association via a Raman process compared to Raman transfer between hyperfine states in atoms is that the matrix elements between the ground states and the excited state are greatly unbalanced. In particular, the matrix element between the molecule state and the excited state, $ \Omega_m $ is much larger than the matrix element between the free atom state and the excited state, $ \Omega_a $. To understand the effect of this inbalance, we consider a 3 level model, and ignore the presence of multiple excited states. The Raman Rabi Rate, $ \Omega_a\Omega_m / 2\Delta$ depends on both matrix elements, but the scattering rate predominantly comes from the final molecular state, $ \Gamma_e \Omega_m^2 / 2\Delta^2$, where $ \Gamma_e $ is the excited state linewidth and $ \Delta $ is the single photon detuning. Assuming the power of each Raman beam is equal, there is an additional factor of 2 in the scattering rate, since the molecular state scatters off of both Raman beams at roughly the same single photon detuning. Thus, the ratio between the Raman Rabi frequency and the scattering rate, $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth. Thus, a small $ \Omega_m/\Omega_a $ ratio is desired to be able to drive this transition coherently. In addition to this fundamental reason, there is also a technical advantage for having a $ \Omega_m/\Omega_a $ ratio to be as small(? or big) as possible. The position of the resonance depends on the laser power predominantly through the AC Stark shift on the molecular state, $ \Omega_m^2 / 2\Delta $. The ratio of the AC Stark shift to the Raman Rabi frequency is $ \Omega_m / \Omega_a $, Thus, the laser needs to be stabilized to better than the inverse of this ratio, $ \Omega_a/\Omega_m $, to fluctuate by less than a linewidth. This becomes technically easier when $ \Omega_m/\Omega_a $ is smaller.
(for our choice of states, this mean intensity fluctuation of less than 1\% or 0.1\%.)

Due to the small size of the molecular wavefunction, the coupling between the ground atomic state and the excited molecular state is approximately proportional to the value of the relative atomic wavefunction at short distance within the molecular potential. In additional to the confinement, this value is related to the interaction between the two atoms. For states with a large scattering length (positive or negative), the phase shift in the relative wavefunction between the atoms can significantly increase the short range wavefunction (add figure?). The increase in the coupling is proportional to (quote/cite Olive's equation?). For our system, among the stable spin combinations, 4422 and 3311 has a small scattering length of $\abs{a}=....$ respectively and 3322 has a large and negative scattering length. (interaction shift $\approx$ binding?) Thus, we use the 3322 spin combination as our initial state and drive to the first bound state for the 3322 spin combination.

% In additional to the final and the excited state, it is also important to select the an initial atomic state in order to improve the coupling.

>>>>

Such ultracold molecular association takes advantage of the much developed cooling and trapping techniques for atoms  as a starting point. To overcome the challenges of small wavefunction overlap and the large release of binding energy of converting atoms to deeply-bound molecules, a  two-step approach has been established to first associate atom pairs into weakly-bound molecules, and then transfer the molecules from this single internal state to a desired rovibrational and electronic state~\cite{Danzl2008, Ni2008,Lang2008, Takekoshi2014, Molony2014, Park2015, Guo2016, Kondov2019, Voges2020}.
So far, all of such association processes  utilized a magnetic Feshbach scattering resonance and have been applied to bialkali molecules. The only exception is Sr$_2$ where narrow linewidth excited states are available and optical association can be driven~\cite{Reinaudi2012,Stellmer2012}. The requirement of a Feshbach resonance to enhance atom-to-molecule wavefunction overlap or the narrow excited state lines limits the generality of the association technique. 
%to more diverse molecular species.

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule using a two-photon optical Raman transfer via an electronic excited state, without the use of a Feshbach resonance nor a narrow excited state. Our scheme is based on a theoretically determined vibrational state of the electronic excited state $c^3\Sigma^+$ that has the best transfer Rabi coupling to photon scattering ratio. To further increase this ratio and reduce technical requirements such as intensity stability, we choose an initial and final state where the matrix elements with the excited state are as balanced as possible. This approach minimizes the reliance on system specific properties and can therefore be more generally applied to creating other molecular species or larger molecules atom-by-atom with full quantum state control. %The choice of the states is general and can be applied to associate other molecular species or large molecules atom-by-atom with full quantum state control. 
(also this result represents full quantum-state control of single [neutral] molecules, the second time ever (first being our NaCs Feshbach molecules)) %This approach minimizes the reliance on system specific properties and can therefore be more generally applied to creating other molecules.
%electronic the best initial hyperfine combination of the atoms to use and best single photon frequency that balances coupling/transfer efficiency and photon scattering. Using these parameters, we demonstrate coherent association of a electronic ground-state molecule from two atoms via an optical Raman transition in an optical tweezer.

@@

%While optical Raman transfers of atoms to molecules utilizing an electronic excited state have been shown previously, the transfers were incoherent  where the resulting molecules were lost~\cite{Wynar2000,Rom2004}.

The challenge of optical Raman transfers of atoms to molecules lies with achieving a high enough Raman Rabi frequency  due to the wavefunction size mismatch with the excited molecular state. While pioneering experiments used weakly bound molecular excited states in the Raman transition to increase the Raman Rabi frequency, this choice of state also increases the scattering during the transfer process to render it incoherent, resulting in molecule loss~\cite{Wynar2000,Rom2004}. 
%Previous experiments used weakly bound molecular excited states in the Raman transition to increase the Raman Rabi frequency. However, this choice of state also increases the scattering during the transfer process for two reasons.
In general, scattering can be reduce by detuning away from resonance. However, we are constrained by the multi-level of molecules. 
Firstly, since for a typical molecular potential the spacing between states
decreases with increasing vibrational levels,
a weakly bound excited molecular state limits the maximum usable detuning and therefore the effectiveness of scattering reduction by increasing the detuning for the Raman transition. Secondly, due to size similarity(?), the weakly bound molecular ground state has a strong coupling with the atomic excited state which contributes significantly to the total scattering rate. The rate of such a scattering process is larger for the weakly bound states, and is proportional to $1/\delta_{thresh}^2$,
where $\delta_{thresh}$ is the detuning from the dissociation threshold. (Should we also mention scattering of the atomic initial state too? This should affect the Raman process as well)
% This further increases the scattering rate for a weakly bond excited state.
Based on these considerations, we calculated both the Raman Rabi frequency and the scattering rate at different detunings (full result is shown in SM). As seen in figure (1....), we found a distinct regime from what have been explored previously, namely, despite a smaller Raman Rabi frequency at deeply bound intermediate states, the Rabi frequency to scattering ratio is much higher as compared to a weakly bound intermediate state. As a result, we use a deeply bound molecular excited state ($v'=0$ of $c^3\Sigma^+$) to drive the Raman transition. (mention the trend, also v=12 will work as similarly as v=0)

%higher Raman Rabi frequency at the same detuning, using a weakly bound intermediate state results in a much higher scattering rate and ultimately a smaller Rabi frequency to scattering rate ratio as compared to deeply bound states. As a result, we use a deeply bound molecular excited state to drive the Raman transition.

In addition to the intermediate state, the choice of the initial and final states are critical for both fundamental and technical reasons. A key difference in molecular association via a Raman process compared to Raman transfer between hyperfine states in atoms is that the matrix elements between the ground states and the excited state are greatly unbalanced. In particular, the matrix element between the molecule state and the excited state, $ \Omega_m $ is much larger than the matrix element between the free atom state and the excited state, $ \Omega_a $. To understand the effect of this imbalance, we consider a 3 level model, and ignore the presence of multiple excited states. The Raman Rabi Rate, $ \Omega_a\Omega_m / 2\Delta$ depends on both matrix elements, but the scattering rate predominantly comes from the molecular state, $ \Gamma_e \Omega_m^2 / 2\Delta^2$, where $ \Gamma_e $ is the excited state linewidth and $ \Delta $ is the single photon detuning. We choose the power of the Raman beams to be equal (this gives the smallest scattering), there is an additional factor of 2 in the scattering rate, since the molecular state scatters off of both Raman beams at roughly the same single photon detuning. Thus, the ratio between the Raman Rabi frequency and the scattering rate, $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth. Thus, a small $ \Omega_m/\Omega_a $ ratio is desired to be able to drive this transition coherently at smaller detunings where the Raman Rabi frequency will be larger. In addition to this fundamental reason, there is also a technical advantage for having a $ \Omega_m/\Omega_a $ ratio to be as small( KW: I think small, but open question whether we should always talk about $\Omega_m/\Omega_a$ or $ \Omega_a/\Omega_m$, KN: maybe $ \Omega_a/\Omega_m$ is better because it sounds more natural that we want to make it large? but either way.) as possible. The position of the resonance depends on the laser power predominantly through the AC Stark shift on the molecular state, $ \Omega_m^2 / 2\Delta $. The ratio of the AC Stark shift to the Raman Rabi frequency is $ \Omega_m / \Omega_a $, Thus, the laser needs to be stabilized to better than the inverse of this ratio, $ \Omega_a/\Omega_m $, to fluctuate by less than a linewidth. This becomes technically easier when $ \Omega_m/\Omega_a $ is smaller.
%(for our choice of states, this mean intensity fluctuation of less than 1\% or 0.1\%. KW: I mentioned this in next paragraph, does it still need to be mentioned here?)

This ratio can be made smaller by increasing the coupling of the atomic state with the excited molecular state. Due to the small size of the molecular wavefunction, the coupling between the ground atomic state and the excited molecular state is approximately proportional to the value of the relative atomic wavefunction at short distance within the molecular potential. In additional to the confinement, this value is related to the interaction between the two atoms. For states with a large scattering length (positive or negative), the phase shift in the relative wavefunction between the atoms can significantly increase the short range wavefunction (add figure?). The increase in the coupling is proportional to (quote/cite Olive's equation?). For our system, among the stable spin combinations, 4422 and 3311 both have small scattering lengths of $\abs{a_{4422}}= 30.4a_0, \abs{a_{3311}} = 13.7a_0 $ respectively and 3322 has a large and negative scattering length of $ \abs{a_{3322}} = 693.8a_0 $ (interaction shift $\approx$ binding?)~\cite{Hood2019}. In addition to the increased atomic coupling, $ \Omega_a $, with the 3322 hyperfine combination, coupled channel calculations show that the most weakly bound molecular state that is predominantly this spin combination also has reduced coupling, $ \Omega_m $, with the excited state when compared with the 4422 and 3311 bound states. Therefore, using an initial 3322 hyperfine combination results in a $ \Omega_m/\Omega_a$ ratio of about 20 instead of a ratio of about 300 with the 4422 or 3311 combinations. This more favorable ratio also relaxes the technical requirement of intensity stability of $0.3 \% $ for the 4422 combination compared to $ 5\% $ for the 3322 combination.  Thus, we use the 3322 spin combination as our initial state and drive to the first bound state for the 3322 spin combination. (We might want to mention that those ratios are for 80 kHz spherical trap if we want to give more details about the coupled channel calculation anywhere...)

% In additional to the final and the excited state, it is also important to select the an initial atomic state in order to improve the coupling.
\begin{figure*}
  \includegraphics[height=4.5cm]{fig2.pdf}
  \caption{
    (A) Geometry and polarization of trap and Raman beam relative to the bias magnetic field.
    The tweezer and Raman beam is focused through an objective to define the location of the
    atoms and molecule.
    We use a bias B field of $B_0=8.8 G$ alone the tweezer polarization
    to define the quantization axis.
    As a result, the atoms experiences predominately $\pi$ polarization from the tweezer.
    (B) Molecule formation pulse sequence. The tweezer initially consists of only up leg power.
    When driving the Raman transition, the up leg power is smoothly ramped down and
    the down leg power ramped up over $10\mu s$ while maintaining the total power of the tweezer.
    This minimizes the heating on the atoms due to power fluctuation while maximizes the time
    with maximum Raman Rabi frequency when the up and down leg powers are equal.
    (C) Raman resonance from atomic state to molecular state, showing Fourier limited linewidth.
    \todo{states, time}
    (B) Rabi oscillation on resonance
    \label{f-raman}}
\end{figure*}

@@
